---
title: "Final Exam on Scorpion Behavior"
author: "Katrina Mercado"
date: "5-26-16"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(mosaic)
require(UsingR)
require(tigerstats)

```

# Introduction

<In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?>

Scorpion Behavior analysis derives from two comma seperated value files we received from a professor studying scorpions. The professor wants to know if scorpions have a sense of smell (or perhaps a nose). In order for us to accurately answer his inquiry, we must ask ourselves the following questions:

Does the presence of odor change the expected outcome on our mearuements?

Does the size of the scorpion change the expected outcome on our measurements?

In the files the professor has given us, these are the measurements we received:

* **scorpion**-a tag indicating which scorpion was emasured
* **behavior** - a factor variable (no sting, dry sting, wet sting,jab)
* **volume**-a numerical measure of venom amount (wet sting)
* **probes**-a numerical count of the number of "hits" applied to get a reaction
* **reaction_time**-a numerical measure of time it took to get a reaction
* **size**-a factor(large, small) of scorpion size
* **roll**-a factor-(control,threat) of absense or prescence of an apparent threat to the scorpion(rat urine)

In this statistical analysis, there were 30 different scorpions tested. Each scorpion was tested twice:
  * 1.) in control( no rat urine present)
  * 2.) in threat (rat urine). 
The order of testing was randomly assigned.

One of the csv files refers to the individual tests run and has 60 observations. The other file is conveniently organized as paired data with observations of each scorpion in both threat and control conditions for conceinence in paired tests.
#### what was asked for
Investigate whether the odors from a predator might influence the defensive behavior of Hadrurus arizonensis, and if the scorpion size plays a role in the type of response.

we see this as two questions

##### does presence of odor change the expected outcome on our measurements?

<for continuous measurements we could ask for center differences (means or medians)

for catagorical explanitory variables (such as count of type of response)
a chi-squared test would be chosen.>

##### does size of the scorpion change the expected outcome on our measurements?

<for continuous measurements we could ask for center differences (means or medians)

for catagorical explanitory variables (such as count of type of response)
a chi-squared test would be chosen.>


# Methods for presence vs absence of odor
In order to analyze the data and be able to understand the research questions we will need to have certain methods for how we plan to display our findings of the data analysis. For graphical results we will use a bar chart in order to clearly display the difference between the behavior of the scorpions when a threat is present or not present.This graph will visually display the difference between the scorpions when looking at their behavior when a threat is present or not.
<Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.>

#Results

<Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.>

## Descriptive Results
In order to understand the comparison between the scorpions when odor is a factor to explain when they would attack or become more aggrevated. We must create a plot in order to compare the reaction time for the scorpions when they feel threatened by a predator and when they do not feel threatened by a predator. We must also produce a t.test in order to show the comparison between the control group as well as the experimental group so that we may see if they have a significant comparison.
### Graphical Descriptive Results

```{r}
url2="scorpion_reaction_rolls.csv"
scorp_data2=read.csv(url2,header=T)
contTable <- xtabs(~ Behavior + roll, data = scorp_data2)
barchart(contTable,
	auto.key = list(
		space = "top",
		title = "roll",
		cex.title = 1,
		columns = 1),
	horizontal = FALSE,
	main = "Behavior",
	xlab = "Roll",
	ylab = "Frequency")
```


The bar graph clearly shows that the scorpions are much more likely to have a wet sting when a threat is present, than in the control group when the threat is not present.The graph also shows that the scorpions would engage in some type of protective instinct whether it be a sting or a jab when the prescence of a threat is detected.The number of scorpions who would have a wet sting when the threat is present is significantly greater than the number of scorpions who do not have a threat present. More scorpions are shown to have a dry sting when there is no threat detected than when there is a threat detected. The overall analysis of this graph would show that when there is a threat present there is more likely to have a wet sting rather than when there is not a threat present which would make the scorpions more likely to have a dry sting.
<Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question>

### Numerical Descriptive Results

```{r}
attach(scorp_data2)
rol=table(roll,Behavior)
rowPerc(rol)
```

During the rowPerc test it clearly shows that the scorpions are more likely to have a wet sting when a threat is present as opposed to when there is no threat present. About 70% of the scorpions will have a wet sting when the threat is present; while those who did not have a threat present only about 7% had a wet sting. When a threat is present almost 17% of the scorpions would jab. In addition to when there is not a threat present 20% of the scorpions would jab. The fact that none of the scorpions would sting or jab when there was a threat present shows that scorpions are extrememly more likely to have some type of protective instinct when the threat is present. This would show that odor plays a huge role when comparing odor and the change in our measurements.


## Inferential Results
```{r}
chisqtestGC(rol,simulate.p.value = TRUE ,graph=TRUE)
```
The chisqtest shows that we would reject the null hypothesis based on the fact that the wet sting was used more frequently with the threat prescence in comparison to when there is no threat present. The pvalue is significantly low which indicates that there is a difference between the two variables compared.
<State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question>

# Methods for different sizes (large vs small)

<Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.>

#Results

<Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.>

## Descriptive Results

### Graphical Descriptive Results

<Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question>

### Numerical Descriptive Results

<Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question>

## Inferential Results

<State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question>


# Discussion and Conclusion for both set of research questions

<Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.>
